---
title: "Productivity estimates for the EAM by Shri Shri Sneha Manda"
output: html_notebook
---

I have a rawdata file with estimated producity rates of each dominant species of benthic foraminifera in the Vermitid reefs of the Levant. In this notebook, I will primarily use dplyr, tidyverse and tidyr packages to clean/summaries/structure the data such that I can incorporate in with the file " Productivity Rates_barplot " that is being used in the R notebook : bf_prod_subtropics.Rmd

```{r}
#setting working directory

setwd("E:/OneDrive - post.bgu.ac.il/Modern Foraminifera Projects/Carbonate Production")

# load data - choose csv file (foram_prod_sm)
sm.data <- read.csv(file.choose())

# The unique() function in R returns a vector, data frame, or array-like object with duplicate elements and rows deleted.

unique(sm.data$foram.type)

str(sm.data)

head(sm.data)
```
```{r}
# To get average productivity rates, I need to summarise the data for each station, e.g: sum of all rows with the entry Akhziv


sm.data %>%
  group_by(Location, foram.type) %>%
  summarize(TotalProduction = sum(carb_prod), na.rm = TRUE)

sm.data %>%
               group_by(Location) %>%
               summarize(TotalProduction = sum(carb_prod), na.rm = TRUE)


sm.data %>%
               group_by(Size.wise.species, Location) %>%
               summarize(TotalProduction = sum(carb_prod), na.rm = TRUE)
```

Okay, I have interesting information here. The carbonate production by foraminifera from the Mediterranean can go as high as 700 g m^-2 yr^-1. TO emphasize, this is MY model,  not Pamela Hallocks's tunover model revised by many authors to give obnoxious results. 

I want to show this on a plot such that foraminiferal productivity rate from the Mediterranean can be compared to those of the global foraminiferal productivity rate. 

```{r}
# loading data file
f.prod <- read.csv(file.choose())

# file name: Productivity Rates_barplot (JNP_SM folder)
# file name: global_foram_cprod_database ## cotains all edits and removal of replicate studies. 

# The unique() function in R returns a vector, data frame, or array-like object with duplicate elements and rows deleted.

unique(f.prod$Oceanic.region)

str(f.prod)
```
I want to introduce 4 more rows to this dataframe such that they contain my data


```{r}
# Adding new rows to dataframe f.prod.small
df.prod <- f.prod %>% 
   add_row(Area = "Akhziv", Oceanic.region = "Mediterranean", worldloc   = "33 n", Latitude = 33, Hemisphere = "n", Estimated.Carbonate.production = 144, Reference = "This study" )
```

```{r}
# creating new data frame

# Create dataframe
df = data.frame(Area=c("Shikmona", "Nahsholim", "Palmachim"),
              Oceanic.region=c("Mediterranean", "Mediterranean", "Mediterranean"), 
              worldloc = c("33 n", "33 n", "32 n"),
              Latitude = c(33, 33, 32), 
              Hemisphere = c("n", "n", "n"), 
              Estimated.Carbonate.production = c(176, 187, 24), 
              Reference = c("This study" , "This study" , "This study" ))

df.prod.sneha <- rbind(df.prod, df)

unique(df.prod.sneha$Oceanic.region)
```

```{r}
# combining different variables

# Let me try to minimize the ocean provinces to see the data better

library(dplyr)

f.prod.small <- df.prod.sneha %>%
  mutate(Realm =case_when(
    Oceanic.region=="Indian Ocean" | Oceanic.region=="Asian" ~ "Indian Ocean",
    Oceanic.region=="Carribbean" | Oceanic.region=="n - atlantic" | Oceanic.region=="s - atlantic"~ "Atlantic", 
    Oceanic.region=="Red Sea" | Oceanic.region=="Persian Gulf" ~ "Middle east", 
    Oceanic.region=="s - pacific" | Oceanic.region=="s—pacific" ~ "South Pacific", 
    Oceanic.region=="n - pacific" ~ "North Pacific",
    Oceanic.region=="Mediterranean" ~ "Mediterranean", 
  ))

```

```{r}

color_realm <- c("#EFF3FF", "#C6DBEF", "#c15b4e","#9ECAE1", "#6BAED6" ,"#3182BD")

plot <- ggplot(f.prod.small, aes(x=reorder(Realm, Estimated.Carbonate.production), y=Estimated.Carbonate.production, fill = Realm, color = Realm)) + 
  geom_violin(outlier.shape = NA) +
  geom_jitter(shape=16,  position=position_jitter(0.2), aes(colour = Realm, alpha = 0.5))+
  scale_fill_grey()+
  stat_summary(fun.y=mean, geom="point", size=2, color="black")+
  ylim(0, 1200)+
  #coord_flip()+
  scale_fill_manual(values = color_realm)+
  scale_colour_manual(values = color_realm)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  theme_bw()+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  ggtitle( bquote(Estimated~CaCO[.(3)]~production~by~reef-dwelling~benthic~foraminifera))+ 
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.title.x=element_blank())+
  labs (y = bquote(grams~m^2~y^-1)) +
  theme(legend.position  = "none")

plot
```
```{r}
#pngfile <- fs::path(knitr::fig_path(),  "15.01.2024_foram_carb_prod_benthic_sm_data_coord_flip.png")
pngfile <- fs::path(knitr::fig_path(),  "23.04.2024_foram_carb_prod_benthic_sm_data_coord_flip.png")
# I'm explicitly calling the device functions so you can see the dimensions 
# used
agg_png(pngfile, width = 30, height = 22, units = "cm", res = 300)
plot(plot)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
#using this chunk of code to extract the colours from the "plot" in R notebook "bf_prod_subtropics" ( R_fill_brewer palette)
ggplot_build(plot)$data
```

```{r}

plot <- ggplot(f.prod.small, aes(x=reorder(Realm, Estimated.Carbonate.production), y=Estimated.Carbonate.production, fill = Realm, color = Realm)) + 
  geom_violin(outlier.shape = NA) +
  geom_jitter(shape=16,  position=position_jitter(0.2), aes(colour = Realm, alpha = 0.5))+
  scale_fill_grey()+
  stat_summary(fun.y=mean, geom="point", size=2, color="black")+
  ylim(0, 1200)+
  #coord_flip()+
  scale_fill_manual(values = color_realm)+
  scale_colour_manual(values = color_realm)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  theme_bw()+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  ggtitle( bquote(Estimated~CaCO[.(3)]~production~by~reef-dwelling~benthic~foraminifera))+ 
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.title.x=element_blank())+
  labs (y = bquote(grams~m^2~y^-1)) +
  theme(legend.position  = "none")

plot
```


```{r}
#plotting results for Israeli MEditerrean by species 

plot2 <- ggplot(sm.data, aes(x=reorder(Size.wise.species, -carb_prod), y=carb_prod, fill = Size.wise.species, color = Size.wise.species)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16,  position=position_jitter(0.2), aes(colour =Size.wise.species , alpha = 0.5))+
  #scale_fill_grey()+
  stat_summary(fun.y=mean, geom="point", size=2, color="black")+
  #ylim(0, 1200)+
  #coord_flip()+
  scale_fill_manual(values = color_realm)+
  scale_colour_manual(values = color_realm)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  theme_bw()+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #ggtitle( bquote(Estimated~CaCO[.(3)]~production~by~reef-dwelling~benthic~foraminifera))+ 
  theme(plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.title.x=element_blank())+
  labs (y = bquote(grams~m^2~y^-1)) +
  theme(legend.position  = "none")

plot2

```

```{r}
color_realm <- c("#EFF3FF", "#C6DBEF", "#c15b4e","#9ECAE1", "#6BAED6" ,"#3182BD")

plot <- ggplot(f.prod.small, aes(x=reorder(Realm, Estimated.Carbonate.production), y=Estimated.Carbonate.production, fill = Realm, color = Realm)) + 
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(shape=16,  position=position_jitter(0.2), aes(colour = Realm, alpha = 0.5))+
  scale_fill_grey()+
  stat_summary(fun.y=mean, geom="point", size=2, color="black")+
  ylim(0, 1000)+
  #coord_flip()+
  scale_fill_manual(values = color_realm)+
  scale_colour_manual(values = color_realm)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #scale_fill_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  theme_bw()+
  #scale_color_discrete_divergingx( palette = "ArmyRose", alpha = 1,rev= TRUE)+
  #ggtitle( bquote(Estimated~CaCO[.(3)]~production~by~reef-dwelling~benthic~foraminifera))+ 
  #theme(plot.title = element_text(hjust = 0.5), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(),      axis.title.x=element_blank())+
  theme(plot.title = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.background = element_blank(), axis.title.x=element_blank())+
  #labs (y = bquote(grams~m^2~y^-1)) +
  theme(legend.position  = "none")

plot
```
```{r}
write.csv(f.prod.small,"Table2_Manda_et_al_2024.csv") 
```

```{r}
#pngfile <- fs::path(knitr::fig_path(),  "15.01.2024_foram_carb_prod_benthic_sm_data_coord_flip.png")
#pngfile <- fs::path(knitr::fig_path(),  "11.08.2024_foram_carb_prod_benthic_sm_data_coord_flip.png")
pngfile <- fs::path(knitr::fig_path(),  "10.02.2025_foram_carb_prod_benthic_sm_data_feb2025_updated_ylim1000_2.png")
# I'm explicitly calling the device functions so you can see the dimensions 
# used
agg_png(pngfile, width = 30, height = 22, units = "cm", res = 300)
plot(plot)
invisible(dev.off())
knitr::include_graphics(pngfile)
```
```{r}
## Method 3: Gives count, mean, standard deviation, standard error of the mean, and confidence interval (default 95%).
  ##   data: a data frame.
  ##   measurevar: the name of a column that contains the variable to be summariezed
  ##   groupvars: a vector containing names of columns that contain grouping variables
  ##   na.rm: a boolean that indicates whether to ignore NA's
  ##   conf.interval: the percent range of the confidence interval (default is 95%)
  
  
  summarySE <- function(data=NULL, measurevar, groupvars=NULL, na.rm=FALSE,
                        conf.interval=.95, .drop=TRUE) {
    install.packages("plyr")
    library(plyr)
    
    # New version of length which can handle NA's: if na.rm==T, don't count them
    length2 <- function (x, na.rm=FALSE) {
      if (na.rm) sum(!is.na(x))
      else       length(x)
    }
    
    # This does the summary. For each group's data frame, return a vector with
    # N, mean, and sd
    datac <- ddply(data, groupvars, .drop=.drop,
                   .fun = function(xx, col) {
                     c(N    = length2(xx[[col]], na.rm=na.rm),
                       mean = mean   (xx[[col]], na.rm=na.rm),
                       sd   = sd     (xx[[col]], na.rm=na.rm)
                     )
                   },
                   measurevar
    )
    
    # Rename the "mean" column    
    datac <- rename(datac, c("mean" = measurevar))
    
    datac$se <- datac$sd / sqrt(datac$N)  # Calculate standard error of the mean
    
    # Confidence interval multiplier for standard error
    # Calculate t-statistic for confidence interval: 
    # e.g., if conf.interval is .95, use .975 (above/below), and use df=N-1
    ciMult <- qt(conf.interval/2 + .5, datac$N-1)
    datac$ci <- datac$se * ciMult
    
    return(datac)
  }
```

```{r}
hist(f.prod.small$Estimated.Carbonate.production, 
     main = "All records of benthic foraminiferal carbonate production",
     xlab = "rate of carbonate production (grams)")
```
```{r}
boxplot(f.prod.small$Estimated.Carbonate.production, 
     main = "All records of benthic foraminiferal carbonate production",
     xlab = "rate of carbonate production (grams)", 
     horizontal = TRUE)
```
```{r}
#doing some stats on this, to nderstand the skweness of the dataset

install.packages("moments")
library(moments)

#calculate skewness
skewness(f.prod.small$Estimated.Carbonate.production)

#calculate kurtosis
kurtosis(f.prod.small$Estimated.Carbonate.production)

#to test the nirmality of data distribution

jarque.test(f.prod.small$Estimated.Carbonate.production)

x <- runif(100)
quantile(f.prod.small$Estimated.Carbonate.production,probs=c(.025,.95))

qts <- quantile(f.prod.small$Estimated.Carbonate.production,probs=c(.05,.95))
hist(f.prod.small$Estimated.Carbonate.production)
abline(v=qts[1],col="red")
abline(v=qts[2],col="red")

qts <- quantile(f.prod.small$Estimated.Carbonate.production,probs=c(.05,.94))
hist(f.prod.small$Estimated.Carbonate.production)
abline(v=qts[1],col="green")
abline(v=qts[2],col="green")
```


```{r}
summary(f.prod.small$Estimated.Carbonate.production)

#calculate range values
range(f.prod.small$Estimated.Carbonate.production, na.rm=TRUE)
```


```{r}
#Let me find out the outliers of this box plot

boxplot.stats(f.prod.small$Estimated.Carbonate.production)$out

#which of these points contribute to the outliers
out <- boxplot.stats(f.prod.small$Estimated.Carbonate.production)$out
out_ind <- which(f.prod.small$Estimated.Carbonate.production %in% c(out))
out_ind
```


```{r}
quartiles <- quantile(f.prod.small$Estimated.Carbonate.production)
print(quartiles)
```

```{r}
hist(f.prod$Estimated.Carbonate.production,
     main = "All records of benthic foraminiferal carbonate production",
     xlab = "rate of carbonate production (grams)")
```
```{r}
#Extact GnBu color details from "Hcl" colourspace

swatchplot(GnBu)

palette2colors(GnBu, reverse = FALSE)
```


```{r}

# View a single RColorBrewer palette by specifying its name
display.brewer.pal(n = 8, name = 'GnBu')

# Hexadecimal color specification 
brewer.pal(n = 8, name = "GnBu")

alldata <- boxplot(f.prod$Estimated.Carbonate.production, 
     main = "All records of benthic foraminiferal carbonate production",
     xlab = "rate of carbonate production (grams CaCO3 per sqm per year)", 
     horizontal = TRUE)

alldata

#which of these points contribute to the outliers
out <- boxplot.stats(f.prod$Estimated.Carbonate.production)$out
out
out_ind <- which(f.prod$Estimated.Carbonate.production %in% c(out))
out_ind

```

```{r}
pngfile <- fs::path(knitr::fig_path(),  "05.03.2025_all_foram_carb_prod.png")
# I'm explicitly calling the device functions so you can see the dimensions 
# used
agg_png(pngfile, width = 22, height = 12, units = "cm", res = 300)
plot(alldata)
invisible(dev.off())
knitr::include_graphics(pngfile)
```

```{r}
summary(f.prod$Estimated.Carbonate.production)
```

```{r}
#calculate skewness
skewness(f.prod$Estimated.Carbonate.production)

#calculate kurtosis
kurtosis(f.prod$Estimated.Carbonate.production)

#to test the nirmality of data distribution

jarque.test(f.prod$Estimated.Carbonate.production)

x <- runif(100)
quantile(f.prod$Estimated.Carbonate.production,probs=c(.025,.95))

qts <- quantile(f.prod$Estimated.Carbonate.production,probs=c(.05,.95))
hist(f.prod$Estimated.Carbonate.production, 
     main = "All records of benthic foraminiferal carbonate production",
     xlab = "rate of carbonate production (grams CaCO3 per sqm per year)")
abline(v=qts[1],col="red")
abline(v=qts[2],col="red")

qts <- quantile(f.prod$Estimated.Carbonate.production,probs=c(.05,.94))
hist(f.prod$Estimated.Carbonate.production,
     main = "All records of benthic foraminiferal carbonate production",
     xlab = "rate of carbonate production (grams CaCO3 per sqm per year)" )
abline(v=qts[1],col="green")
abline(v=qts[2],col="green")
```
```{r}
#writing a table

write.csv(f.prod.small, "/Users/snehamanda/Library/CloudStorage/OneDrive-post.bgu.ac.il/Modern Foraminifera Projects/Carbonate Production/Feb 2025/Suppl.Table2.csv", fileEncoding = "UTf-8")

```

